---
title: "Star-Wars-EP4-Script-Text-Analysis"
author: "maxsharman "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Loading Packages


```{r}
library(textstem)
library(tidyverse)
library(stringr)
library(tidytext)
library(stringr)
library(knitr)
library(ggplot2)
library(wordcloud)
```

#### Reading Star Wars Episode 4 Script into R

```{r}
script <- read.delim('SW_EpisodeIV.txt') 
```

#### Exploring the Text File

```{r}
str(script) 
```
Text Analysis On The Whole Script
===

#### Extracting Character Names

```{r}
Characters <- script %>% 
  mutate(character.names = str_extract_all(character.dialogue, "\\b[A-Z][A-Z]+\\b")) %>% 
  filter(lengths(character.names) > 0) %>% 
  mutate(character.names = str_replace_all(character.names, ",", ""))
```
#### Finding Top 5 Character Name Occurances and thier Frequency Counts

```{r}
character_dialogue_counts <- table(unlist(Characters$character.names))

character_dialogue_counts <- sort(character_dialogue_counts, decreasing = TRUE)

character_counts_df <- as.data.frame(character_dialogue_counts)

top_characters <- character_counts_df %>% head(5)

kable(top_characters)
```
#### Plotting the Top 5 Occuring Characters And their Dialogue Frequency

```{r}
Characters <- ggplot(top_characters, aes(Var1, Freq, fill = Var1)) +
geom_bar(stat = 'identity', color = "Black") + ggtitle('Star Wars Ep 4
Top 5 Character Dialogue Frequency') + xlab("Characters") +
ylab("Frequency") + labs(fill = "Characters") + theme_minimal() +
scale_fill_brewer(palette = "Blues")

print(Characters)
```


```{r}
character_names <- c(
  "luke", "han", "threepio", "ben", "leia", "vader", "red leader", "biggs", "tarkin", "owen",
  "trooper", "gold leader", "wedge", "officer", "red ten")

Tokenized_script <- script %>%
  mutate(text = as.character(character.dialogue)) %>%
  unnest_tokens(word, text) %>% 
  anti_join(get_stopwords()) %>% 
  filter(!str_detect(word, "\\d+")) %>% 
  filter(!word %in% character_names)

```

#### Counting word occurrences
```{r}
Tokenized_script <- Tokenized_script %>% mutate(lemmatized_text = lemmatize_words(word))
Tokenized_script <- drop(Tokenized_script['lemmatized_text'])

word_frequency <- Tokenized_script %>% 
  count(lemmatized_text, sort = TRUE)
```

#### Sorting for the top 10 most frequent words
```{r}
top_10_words <- word_frequency %>% head(10)
top_10_words <- top_10_words %>% rename(Frequency = n)
```

#### Visual Bar chart for the top 10 most frequent words

```{r}
Word_occurrences <- ggplot(top_10_words, aes(lemmatized_text, Frequency, fill = Frequency)) +
  geom_bar(stat = 'identity', color = "Black") +
  ggtitle('Top 10 Most Frequent Words') +
  xlab("word") +
  ylab("Frequency") +
  theme_minimal() +
  scale_fill_gradient(low = "lightblue", high = "darkblue")
  
print(Word_occurrences)

wordcloud(words = word_frequency$lemmatized_text, freq = word_frequency$n, min.freq = 10,
          random.order = FALSE, colors = brewer.pal(8, "BrBG"))
```

#### Script Sentiment counts

```{r}
Sentiment_count <- Tokenized_script %>% 
  inner_join(get_sentiments(), c("lemmatized_text" = "word")) %>% 
  count(sentiment)

kable(Sentiment_count)
```

#### Visual Bar Chart For the Sentiment Analysis

```{r}
Sentiment_count <- ggplot(Sentiment_count, aes(sentiment, n, fill = sentiment)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = n), vjust = -0.3, color = "Black") +
  ggtitle("Star Wars Ep4 Sentiment Analysis") +
  xlab("Sentiments") +
  ylab("Counts") +
  theme_minimal() +
  scale_fill_brewer(palette = "Blues")

print(Sentiment_count)
```


Text Analysis By Character
===

```{r}
Luke_lines <- script %>% 
  filter(grepl("LUKE", character.dialogue))

Vader_lines <- script %>% 
  filter(grepl("VADER", character.dialogue))


Threepio_lines <- script %>% 
  filter(grepl("THREEPIO", character.dialogue))

```

#### LUKE
```{r}

Tokenized_Luke_lines <- Luke_lines %>%
  mutate(text = as.character(character.dialogue)) %>%
  unnest_tokens(word, text) %>% 
  anti_join(get_stopwords()) %>% 
  filter(!str_detect(word, "\\d+")) %>% 
  filter(!word %in% character_names)


Tokenized_Luke_lines <- Tokenized_Luke_lines %>% mutate(lemmatized_text = lemmatize_words(word))
Tokenized_Luke_lines <- drop(Tokenized_Luke_lines['lemmatized_text'])

word_frequency_Luke_lines <- Tokenized_Luke_lines %>% 
  count(lemmatized_text, sort = TRUE)

top_10_words_Luke_lines <- word_frequency_Luke_lines %>% head(10)
top_10_words_Luke_lines <- top_10_words_Luke_lines %>% rename(Frequency = n)
```

#### VADER
```{r}
Tokenized_Vader_lines <- Vader_lines %>%
  mutate(text = as.character(character.dialogue)) %>%
  unnest_tokens(word, text) %>% 
  anti_join(get_stopwords()) %>% 
  filter(!str_detect(word, "\\d+")) %>% 
  filter(!word %in% character_names)

Tokenized_Vader_lines <- Tokenized_Vader_lines %>% mutate(lemmatized_text = lemmatize_words(word))
Tokenized_Vader_lines <- drop(Tokenized_Vader_lines['lemmatized_text'])

word_frequency_Vader_lines <- Tokenized_Vader_lines %>% 
  count(lemmatized_text, sort = TRUE)

top_10_words_Vader_lines <- word_frequency_Vader_lines %>% head(10)
top_10_words_Vader_lines <- top_10_words_Vader_lines %>% rename(Frequency = n)
```

#### THREEPIO
```{r}
Tokenized_Threepio_lines <- Threepio_lines %>%
  mutate(text = as.character(character.dialogue)) %>%
  unnest_tokens(word, text) %>% 
  anti_join(get_stopwords()) %>% 
  filter(!str_detect(word, "\\d+")) %>% 
  filter(!word %in% character_names)


Tokenized_Threepio_lines <- Tokenized_Threepio_lines %>% mutate(lemmatized_text = lemmatize_words(word))
Tokenized_Threepio_lines <- drop(Tokenized_Threepio_lines['lemmatized_text'])

word_frequency_Threepio_lines <- Tokenized_Threepio_lines %>% 
  count(lemmatized_text, sort = TRUE)

top_10_words_Threepio_lines <- word_frequency_Threepio_lines %>% head(10)
top_10_words_Threepio_lines <- top_10_words_Threepio_lines %>% rename(Frequency = n)
```

#### Most Frequent Words By Character
```{r}
 ggplot(top_10_words_Luke_lines, aes(lemmatized_text, Frequency, fill = Frequency)) +
  geom_bar(stat = 'identity', color = "Black") +
  ggtitle('LUKE Top 10 Most Frequent Words') +
  xlab("word") +
  ylab("Frequency") +
  theme_minimal() +
  scale_fill_gradient(low = "lightblue", high = "darkblue")

wordcloud(words = word_frequency_Luke_lines$lemmatized_text, freq = word_frequency_Luke_lines$n, min.freq = 3,
          random.order = FALSE, colors = c('steelblue', 'darkblue', 'navy', 'midnightblue'))
  
```

```{r}
ggplot(top_10_words_Vader_lines, aes(lemmatized_text, Frequency, fill = Frequency)) +
  geom_bar(stat = 'identity', color = "Black") +
  ggtitle('VADER Top 10 Most Frequent Words') +
  xlab("word") +
  ylab("Frequency") +
  theme_minimal() +
  scale_fill_gradient(low = "red", high = "darkred")

wordcloud(words = word_frequency_Vader_lines$lemmatized_text, freq = word_frequency_Vader_lines$n, min.freq = 2,
          random.order = FALSE,  colors = c('firebrick', 'darkred', 'indianred'))
```

```{r}
 ggplot(top_10_words_Threepio_lines, aes(lemmatized_text, Frequency, fill = Frequency)) +
  geom_bar(stat = 'identity', color = "Black") +
  ggtitle('THREEPIO Top 10 Most Frequent Words') +
  xlab("word") +
  ylab("Frequency") +
  theme_minimal() +
  scale_fill_gradient(low = "lightyellow", high = "yellow")

wordcloud(words = word_frequency_Threepio_lines$lemmatized_text, freq = word_frequency_Threepio_lines$n, min.freq = 3,
          random.order = FALSE, colors = c('darkorange', 'gold', 'darkgoldenrod'))

```



```{r}
VADER_Sentiment_count <- Tokenized_Vader_lines %>% 
  inner_join(get_sentiments(), c("lemmatized_text" = "word")) %>% 
  count(sentiment)

LUKE_Sentiment_count <- Tokenized_Luke_lines %>% 
  inner_join(get_sentiments(), c("lemmatized_text" = "word")) %>% 
  count(sentiment)

THREEPIO_Sentiment_count <- Tokenized_Threepio_lines %>% 
  inner_join(get_sentiments(), c("lemmatized_text" = "word")) %>% 
  count(sentiment)
```


#### Sentiment Count By Character

```{r}
LUKE_Sentiment_count_plot <- ggplot(LUKE_Sentiment_count, aes(sentiment, n, fill = sentiment)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = n), vjust = -0.3, color = "Black") +
  ggtitle("LUKE Star Wars Ep4 Sentiment Analysis") +
  xlab("Sentiments") +
  ylab("Counts") +
  theme_minimal() +
  scale_fill_manual(values = c("skyblue", 'blue'))

print(LUKE_Sentiment_count_plot)
```
```{r}
VADER_Sentiment_count_plot <- ggplot(VADER_Sentiment_count, aes(sentiment, n, fill = sentiment)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = n), vjust = -0.3, color = "Black") +
  ggtitle("VADER Star Wars Ep4 Sentiment Analysis") +
  xlab("Sentiments") +
  ylab("Counts") +
  theme_minimal() +
  scale_fill_manual(values = c("red", 'darkred'))

print(VADER_Sentiment_count_plot)
```



```{r}
THREEPIO_Sentiment_count_plot <- ggplot(THREEPIO_Sentiment_count, aes(sentiment, n, fill = sentiment)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = n), vjust = -0.3, color = "Black") +
  ggtitle("THREEPIO Star Wars Ep4 Sentiment Analysis") +
  xlab("Sentiments") +
  ylab("Counts") +
  theme_minimal() +
  scale_fill_manual(values = c("yellow", 'lightyellow'))

print(THREEPIO_Sentiment_count_plot)
```


















